---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

# uklandusedata

<!-- badges: start -->
<!-- badges: end -->

The goal of uklandusedata is to show how to get open land data for mapping with open source software such as R and QGIS.

It is a work in progress and does not aim to be comprehensive, it's a big topic!

```{r}
library(sf)
library(tmap)
tmap_mode("view")
```


# Choose region of interest

We will get data for the following region (should work for any region in the UK):

```{r}
region = osmdata::getbb(place_name = "leeds", format_out = "sf_polygon")
qtm(region$polygon$geometry[1])
region = region$polygon$geometry[1]
region_27700 = st_transform(region, 27700)
```


# Woodland grant scheme and wood-pasture and parkland inventory data

To download data from the woodland grant and wood-pasture inventory datasets run the following code.

```{r get-woodland-data, eval=FALSE}
# source("r-code/get-estates-data.R")

if(!file.exists("English_Woodland_Grant_Scheme.Rds")) {
  download.file("https://github.com/Robinlovelace/uklandusedata/releases/download/0.0.1/English_Woodland_Grant_Scheme.Rds", "English_Woodland_Grant_Scheme.Rds")
}

w = readRDS("English_Woodland_Grant_Scheme.Rds")
w_region = w[region_27700, ]
qtm(w_region)
saveRDS(w_region, "w_leeds.Rds")
sf::write_sf(w_region, "w_leeds.gpkg")
piggyback::pb_upload("w_leeds.Rds")
piggyback::pb_upload("w_leeds.gpkg")
```

To get the data saved in the previous step for a specific region you can do:

```{r rmap region}
download.file(
  "https://github.com/Robinlovelace/uklandusedata/releases/download/0.0.1/w_leeds.gpkg",
  "w_leeds.gpkg"
)
w_region = sf::read_sf("w_leeds.gpkg")
head(w_region)
plot(w_region)
```

And to make an interactive map, try:

```{r}
tm_shape(w_region) +
  tm_polygons("CurrStat") +
  tm_basemap(leaflet::providers$CartoDB.Positron)
```

## Data from OSM

We can also get data from OpenStreetMap.
A good way to do that is via the QuickOSM plugin in QGIS.

From that the following dataset was generated:

```{r, eval=FALSE, echo=FALSE}
piggyback::pb_upload("landuse_farm_polygons.gpkg", overwrite = T)
```


```{r landuse}
landuse_farm_polygons = sf::read_sf("landuse_farm_polygons.gpkg")

```

